```{r, child="price_adj_setup_shared.qmd"}
```

# Estimation of Price Deflators

To demonstrate how the proposed deflators are estimated in R, we use the 2019 Poland Household Budget Survey (HBS). In this analysis, we aim to estimate spatial price deflators at the NUTS 2 level (regions, *województwa*), which is the lowest NUTS level available in the HBS dataset accessible through the World Bank’s *datalibweb* database.


## Traditional Price deflator with extension: Paasche Index
This section covers two approaches for estimating spatial deflators. The first approach uses the Paasche price index, beginning with the calculation of the food Paasche index and incorporating rent to extend the coverage to non-food expenditures, even in the absence of direct price data.

### Conceptual Overview of the Paasche-Based Tradditional Approach
The following sections present the construction of the food Paasche index and an extended version that incorporates rent to partially capture non-food price variation.


#### Food Paasche Index

The food Paasche is defined as follow:

$$
P_h^{\text{food}} = \left( \sum_{j=1}^F w_j^h \left( \frac{p_j^0}{p_j^h} \right) \right)^{-1}
$$ {#eq-fpaasche}

where $w_j^h$ is the budget share for food item $j$ of household $h$, $F$ is the number of food items considered, and $p_j^h$ and $p_j^0$ are the corresponding market prices of item $j$ for household $h$ and for the reference household $0$, respectively.


#### Food + Rent Paasche index

To increasing the coverage of the deflator, an additional expenditure component, particularly housing expenses is added.  We consider the following index:

$$
P_h^{\text{(food+rent)}} = \left[ \sum_{j=1}^F w_h^j \frac{E[p_h^j]}{p_h^j} + w_h^{\text{rent}} \frac{E_0[x_h^{\text{rent}}]}{E_g[x_h^{\text{rent}}]} \right]^{-1}
$$ {#eq-frpaasche}

where $w_h^j = \frac{x_h^j}{x_h^{\text{food}} + x_h^{\text{rent}}}; w_h^{\text{rent}} = \frac{x_h^{\text{rent}}}{x_h^{\text{food}} + x_h^{\text{rent}}}$

are the budget shares on food item *j* and on housing, respectively, and $E_g[x_h^{\text{rent}}]$ is the average rent for households belonging to group $g$ , while $E_0[x_h^{\text{rent}}]$ is the mean of the group-level rents $E_g[x_h^{\text{rent}}]$. Equation (@eq-frpaasche) is clearly an improvement, in terms of coverage, with respect to equation (@eq-fpaasche).


### Data sets used for the analysis {#sec-dataused}

After constructing the consumption aggregate from the HBS data, the following processed datasets are used to estimate the price deflators. All datasets are saved in `.RDS` format. Each dataset is briefly described below.

```{r, echo = FALSE}
# Processed output data sets saved in data_in:
idata <- tribble(
  ~File, ~Description,
  "hh_info.rds", "HH information (location,etc)",
  "cpi.rds", "Quarterly national price survey",
  "ca_nominal_t2.rds", "Consumption aggregate data with paid and imputed rents"
)
```
```{r  tbl-idata, echo = FALSE, results='asis'}
kable(idata, caption = "Processed HBS data to be used in this chapter")
```


### Preparation of food expenditure data and calculation of food Paasche index
This code creates a dataset containing input for budget shares at hh level that are necessary for food Paasche and food+rent Paasche exteded indexes. 


```{r}
### FOOD EXPENDITURE data ###

# Load and browse food item data
food_item <- readRDS(file.path(data_in, "food_item.rds"))
glimpse(food_item)

# Load and browse quarterly CPI data
cpi <- readRDS(file.path(data_in, "cpi.rds"))
glimpse(cpi)

#Add quarterly rebased CPI
food<- mg(food_item,cpi,key=c("quarter"), drop_merge = TRUE)

# Compute real food expenditure
food <-  food %>%
  mutate(foodexp = cons / cpi,
         foodexp = foodexp / hhsize)
food_expenditure_item <- food 
rm(food)

# Save food expenditure at item level
saveRDS(food_expenditure_item, file.path(data_temp, "food_expenditure_item.rds"))

# Compute food expenditure at household level
food_expenditure <- food_expenditure_item %>%
  group_by(hhid) %>%
  summarise(foodexp = sum(foodexp, na.rm = TRUE), 
            urban = first(urban),
            location = first(location), 
            region = first(region), 
            nuts2 = first(nuts2), 
            mregion = first(mregion))

saveRDS(food_expenditure, file.path(data_temp, "food_expenditure.rds"))

### HOUSING EXPENDITURE ###

# Load and browse consumption aggregate data with rents paid
ca_nominal_t2 <- readRDS(file.path(data_in, "ca_nominal_t2.rds"))
glimpse(ca_nominal_t2)
housing_expenditure <- ca_nominal_t2 %>%
  select(hhid, quarter, ca_rent)

# Load and browse household information data
hh_info <- readRDS(file.path(data_in, "hh_info.rds"))
glimpse(hh_info)

# Merge datasets
housing_expenditure <- mg(housing_expenditure,cpi,key=c("quarter"), drop_merge = TRUE)
housing_expenditure <- mg(housing_expenditure, hh_info, key=c("hhid"), drop_merge = TRUE)

#Calculate monthly food expenditure per capita with price adjusted quarterly
housing_expenditure <- housing_expenditure %>%
  mutate(rentexp = ca_rent / cpi /12/ hhsize)

#Save rent and housing data.
saveRDS(housing_expenditure, file.path(data_temp, "housing_expenditure.rds"))


# Merge food and housing expenditure data
budget_shares <- mg(food_expenditure, housing_expenditure, key=c("hhid"), drop_merge = TRUE)

#Rename variables and calculate the total household food and food & rent total. 
budget_shares <- budget_shares %>%
  rename(expenditure_food = foodexp, expenditure_rent = rentexp)  %>%
  mutate(expenditure_hh_fr = rowSums(select(., expenditure_food, expenditure_rent), na.rm = TRUE)) %>%
  mutate(expenditure_hh_f = expenditure_food)

#Adding variable labels and keep only variables to be used. 
var_label(budget_shares$expenditure_hh_f) <-  "hh expenditure on food"
var_label(budget_shares$expenditure_hh_fr) <- "hh expenditure on food and rent"
budget_shares_input <- budget_shares %>%
  select(hhid,urban,location, region, nuts2, mregion, starts_with(c("expenditure_")), hhsize)

#Descriptive stas of expenditure variables. 
summary(budget_shares_input[, c("expenditure_hh_f", "expenditure_hh_fr", "expenditure_rent")])

#Save the budget shares file to be worked below. 
write_rds(budget_shares_input, file.path(data_temp, "budget_shares_input.rds"), compress = "gz")

```


### Extracting the Elements for Paasche Index Calculation 

#### Food
```{r}
# Load food item data
food_expenditure_item <- readRDS(file.path(data_temp, "food_expenditure_item.rds"))

# Compute mean prices
food_expenditure_item <- food_expenditure_item %>%
  mutate(foodq = cons_q,
         up = foodexp / foodq)

# Compute mean price at national level
up0_mean <- food_expenditure_item %>%
  mutate(up = as.numeric(up)) %>%
  group_by(code) %>%
  summarise(up0_mean = mean(up, na.rm = TRUE))

up0_p50 <- food_expenditure_item %>%
  mutate(up = as.numeric(up)) %>%
  group_by(code) %>%
  summarise(up0_p50 = median(up, na.rm=TRUE))

# Merge price data
food_expenditure_item <- mg(food_expenditure_item, up0_mean, key=c("code"))
if (any(food_expenditure_item$merge != 3)) {
  stop("Check: You have unmatched bservations.")
}
## Make sure if dropping merge=2 is OK if exists.
food_expenditure_item <- food_expenditure_item %>%
  filter(merge!=2) %>% 
  select(-merge) 

## Make sure if dropping merge=2 is OK if exists.
food_expenditure_item <- mg(food_expenditure_item, up0_p50, key=c("code")) %>%
  filter(merge!=2) %>% 
  select(-merge) 

## Make sure if dropping merge=2 is OK if exists.
food_expenditure_item <- mg(food_expenditure_item, budget_shares_input, key=c("hhid")) %>%
  filter(merge==3) %>% 
  select(-merge) 

# Generate budget shares of each food item used Paasche Indexes
food_expenditure_item <- food_expenditure_item %>%
  mutate(w_f_f = foodexp / expenditure_hh_f,
         w_f_fr = foodexp / expenditure_hh_fr)

var_label(food_expenditure_item$w_f_f) <-   "Food item budget share of food paasche"
var_label(food_expenditure_item$w_f_fr) <-  "Food item budget share of extended paasche (food + rent)"

# Check food shares 
check <- food_expenditure_item %>%
  group_by(hhid) %>%
  mutate(check_f = sum(w_f_f, na.rm = TRUE)) %>%
  ungroup()
summary(check$check_f)

# Check if all check_f values round to 1 (equivalent to Stata `assert round(check_f) == 1`)
if (!all(dplyr::near(check$check_f, 1, tol = 0.01))) {
  stop("Not all values of check_f are approximately 1 within a tolerance of 0.01")
}      


# Robustify the denominator of the food paasche (up)	

# Ensure `up` is numeric
food_expenditure_item <- food_expenditure_item %>%
  mutate(up = as.numeric(up))

# Compute `up_mean` and `up_p50` at `(code, nuts2)` level
up_stats <- food_expenditure_item %>%
  group_by(code, nuts2) %>%
  summarise(
    up_mean = mean(up, na.rm = TRUE),
    up_p50 = median(up, na.rm = TRUE),
    .groups = "drop"
  )

# Merge aggregated values back into the full dataset
food_expenditure_item <- mg(food_expenditure_item, up_stats, key=c("code", "nuts2"))

food_expenditure_item <- food_expenditure_item %>%
  mutate(
    check = up0_mean / up_mean,  # Generate check variable
    inve_paasche_food = w_f_f * (up0_mean / up_mean),
    inve_paasche_food_fr = w_f_fr * (up0_mean / up_mean),
    inve_paasche_food_p50 = w_f_f * (up0_p50 / up_p50),  # Now properly sized
    inve_paasche_food_fr_p50 = w_f_fr * (up0_p50 / up_p50)  # Now properly sized
  )


summary(food_expenditure_item$check)
check2<- food_expenditure_item %>%
  group_by(hhid) %>%
  mutate(check2 = sum(w_f_f, na.rm = TRUE)) %>%
  ungroup()
# Assert: Check if all check2 values round to 1 (equivalent to Stata `assert round(check_f) == 1`)
if (!all(dplyr::near(check2$check2, 1, tol = 0.01))) {
  stop("Not all values of check2 are approximately 1 within a tolerance of 0.01")
}            

food_expenditure_item <- food_expenditure_item %>%
  select(-check)

# Aggregate by household
food_file <- food_expenditure_item %>%
  group_by(hhid) %>%
  summarise(
    across(starts_with("inve_paasche"), ~ sum(., na.rm = TRUE)),  # Sum all inve_paasche* variables
    across(starts_with("w_f_"), ~ sum(., na.rm = TRUE))  # Sum all w_f_* variables
  )

label(food_file$inve_paasche_food) <- 	"Food Paasche - HH level"
label(food_file$inve_paasche_food_fr) <- 	"Food innput of extended paasche (food+rent)"
label(food_file$inve_paasche_food_p50) <-     "Food Paasche (p50)- HH level"
label(food_file$inve_paasche_food_fr_p50) <-   "Food input of extended paasche (p50, food+rent)"

summary(food_file[, c("inve_paasche_food","inve_paasche_food_fr", "inve_paasche_food_p50",
                      "inve_paasche_food_fr_p50","w_f_f","w_f_fr")]) ## same as stata 

write_rds(food_file, file.path(data_temp, "food_file.rds"), compress = "gz")
```


#### Non-food (housing)
```{r}
# Load housing data
housing_expenditure <- readRDS(file.path(data_temp, "housing_expenditure.rds")) %>%
  rename(rent_expenditure = rentexp)

# Compute mean rent at national and regional levels
mrent0 <- mean(housing_expenditure$rent_expenditure, na.rm = TRUE)
mrent <- housing_expenditure %>%
  group_by(nuts2) %>%
  summarise(mrent = mean(rent_expenditure, na.rm = TRUE))

housing_expenditure <- mg(housing_expenditure, mrent, key=c("nuts2"))
housing_expenditure <- housing_expenditure %>%
  mutate(mrent0 = if_else(is.na(mrent0), mean(mrent0, na.rm = TRUE), mrent0)) %>%
  select(-merge)

summary(housing_expenditure[, c("mrent0", "mrent")]) ## same as stata 

budget_shares_input <- readRDS(file.path(data_temp, "budget_shares_input.rds"))

# Bring in input for computing budget shares
housing_expenditure <- mg(housing_expenditure, budget_shares_input, key=c("hhid"))

rent_file <- housing_expenditure %>%
  filter(merge==3) %>%
  select(-merge) %>%
  mutate(
    w_r_fr = rent_expenditure / expenditure_hh_fr,  # Compute rent budget share
    inve_paasche_rent_fr = w_r_fr * (mrent0 / mrent)  # Compute Paasche index for rent
  ) %>%
  select(hhid, urban, location, region, nuts2, mregion, w_r_fr, inve_paasche_rent_fr)
label(rent_file$inve_paasche_rent_fr) <- "Rent input of extended paasche (food+rent)"

write_rds(rent_file, file.path(data_temp, "rent_file.rds"), compress = "gz")
```

#### Merging food and non-food(housing) expenditure datasets
```{r}
# Merge food and housing files
paasche_extended <- mg(food_file, rent_file, key=c("hhid"), drop_merge = TRUE)

# Check on shares - Food shares have been checked above
paasche_extended <- paasche_extended %>%
  mutate(check_fr = rowSums(select(., w_f_fr, w_r_fr), na.rm = TRUE))
summary(paasche_extended$check_fr)
stopifnot(all(round(paasche_extended$check_fr, 2) == 1))
paasche_extended <- paasche_extended %>%
  select(-check_fr)
```

### Calculation of Paasche price indexes

```{r}
#Calculation of Food SPI & Extended SPI: food + rent
paasche_extended <- paasche_extended %>%
  mutate(spi_food = 1 / inve_paasche_food,
         spi_food_p50 = 1 / inve_paasche_food_p50,
         spi_food_rent = 1 / (inve_paasche_food_fr + inve_paasche_rent_fr),
         spi_food_rent_p50 = 1 / (inve_paasche_food_fr_p50 + inve_paasche_rent_fr))

label(paasche_extended$spi_food) <- "Paasche Food Index"
label(paasche_extended$spi_food_p50) <- "Paasche Food Index (p50)"
label(paasche_extended$spi_food_rent) <- "Paasche index (food+rent)"
label(paasche_extended$spi_food_rent_p50) <- "Paasche index (p50, food+rent)"

summary(paasche_extended[, c("spi_food", "spi_food_p50", "spi_food_rent", "spi_food_rent_p50")]) ## same as stata 

paasche_extended <- paasche_extended %>%
  select(hhid, urban, location, region, nuts2, mregion, 
         w_f_f, w_f_fr, w_r_fr, 
         inve_paasche_food, inve_paasche_food_p50, 
         inve_paasche_food_fr, inve_paasche_food_fr_p50, 
         inve_paasche_rent_fr, 
         spi_food, spi_food_p50, spi_food_rent, spi_food_rent_p50, everything()) %>%
  arrange(urban, location, region, nuts2, mregion, hhid)

write_rds(paasche_extended, file.path(data_temp, "paasche_extended.rds"), compress = "gz")

```

### Tables for food and extended SPI

```{r}
# Mean spi_food and spi_food_rent at the NUTS2 level
nuts2_summary <- paasche_extended  %>%
  group_by(nuts2) %>%
  summarise(
    mean_spi_food = mean(spi_food, na.rm = TRUE),
    mean_spi_food_rent = mean(spi_food_rent, na.rm = TRUE),
    mean_spi_food_p50 = mean(spi_food_p50, na.rm = TRUE),
    mean_spi_food_rent_p50 = mean(spi_food_rent_p50, na.rm = TRUE)
  )

print(nuts2_summary)

# Mean spi_food and spi_food_rent at the macro region level
mregion_summary <- paasche_extended  %>%
  group_by(mregion) %>%
  summarise(
    mean_spi_food = mean(spi_food, na.rm = TRUE),
    mean_spi_food_rent = mean(spi_food_rent, na.rm = TRUE)
  )

print(mregion_summary)
```

## Full-coverage Price Index

### Conceptual Overview of the Full-coverage Price Index method 
@amendola2024 and @amendola2025 proposed a novel strategy for generating a full-coverage Paasche spatial deflator. The central result of their approach can be expressed as follows:

$$
P_h^{\text{(full-coverage)}} = \left[ w_h^{\text{food}} \left( P_h^{\text{food}} \right)^{-1} + \left(1 - w_h^{\text{food}} \right) \left( P_h^{\text{nfood}} \right)^{-1} \right]^{-1} 
$${#eq-full-coverage}

where $w_h^{\text{food}}$ is the food budget share of household $h$, and $P_h^{\text{food}}$ is the household-level food Paasche index defined in @eq-fpaasche.

$$
P_h^{\text{nfood}} = \frac{1}{1 - w_z} \, \widetilde{\text{SPI}}_g - \frac{w_z}{1 - w_z} \, \text{SPI}_g^{\text{food}}
\quad 
$${#eq-ph-nfood}

where ${SPI}_g^{food}$ is a Paasche spatial food price index for group $g$ – typically a district, province or region, $\widetilde{SPI_g} =  \frac{z_g}{E[z_g]}$ is an implicit deflator defined by the ratio between the group-level absolute poverty line $z_g$ and the average of the $G$ absolute poverty lines, and $w_z$ is the average food budget share of the bottom deciles of the distribution of per capita expenditure. The idea behind this method is that the ratio between absolute poverty lines and their national average can be interpreted as the ratio between the minimum expenditures needed to achieve a fixed minimum level of utility. The reference poverty lines must therefore be absolute, not relative, and must be correctly identified, that is, utility consistent.

The calculation of the CBN poverty lines follows the well-established and widely-used method described in ravallion1998 ravallion2016. As for the calculation of non-food allocations, the upper-bound CBN lines, as opposed to the lower-bound CBN lines, must be used, ensuring that the ratio between the food poverty line and the total poverty line reflects the consumption patterns of poor households. ^[Additionally, as shown in Equation (1) in @menyhert2024, the non-food allowance in the CBN-lower bound line is always, by construction, lower than the food poverty line. This is a constraint that appears rather inappropriate in the context of middle- to high-income economies, where the food budget share is relatively low.]


### Data Sets Used for the Analysis
The same table presented in @tbl-idata is shown again here, with one additional variable required for the analysis in this section.

```{r, echo = FALSE, results='asis'}
kable(idata, caption = "Processed HBS data to be used in this chapter")
```



#### Preparation of Calorie Conversion Table

A calorie conversion table is required to convert food quantities consumed into calories. In this analysis, the authors primarily used publicly available sources such as [https://www.nutritionix.com/](https://www.nutritionix.com/), though additional sources were also consulted. Given the country-specific nature of food consumption, input from local experts would be valuable for finalizing the calorie conversion table.

The finalized conversion table (in Excel format) is saved in the `data_in` folder.

```{r, echo = FALSE}
# Auxiliary calorie table input data:
auxiliary_idata <- tribble(
  ~File, ~Description,
  "kcal_conversion_factors.xlsx", "Prepared calorie table",
)
```

```{r tbl-caltab_xlsx, echo = FALSE, results='asis'}
kable(auxiliary_idata, caption = "Processed HBS data to be used in this chapter")
```


#### Combining Calorie Conversion with Food Item Data

In this step, the calories consumed for each food item are estimated using the Calorie Conversion Table. To control for outliers, households whose per capita calorie consumption falls in the bottom or top 1 percent of the sample are excluded.

```{r}
# Load food calorie conversion factor table in Excel
kcal_conversion_factors <- read_excel(file.path(data_in,"kcal_conversion_factors.xlsx"))

# Clean up the data
## Need to convert to numeric from character
kcal_conversion_factors <- kcal_conversion_factors %>%
  select(-Note)
kcal_conversion_factors$kcal100g <- as.numeric(kcal_conversion_factors$kcal100g)    

# Load food item data and merge with calorie conversion factors
food_item<- readRDS(file.path(data_in, "food_item.rds"))

foodkcal <- mg(food_item, kcal_conversion_factors, key=c("code"))
foodkcal <- foodkcal %>%    
  select(-merge)

#Count if we have missing in kcal100g (calorie conversion) to confirm zero.
sum(is.na(foodkcal$kcal100g))

# Generate kcal (Daily kilocalories consumed per capita) variable for each item
foodkcal$kcal <- foodkcal$cons_q * (foodkcal$edible / 100) * (foodkcal$kcal100g / 100)
# Check var kcal
skim(foodkcal$kcal)%>%
  yank("numeric")


# Tabulate items with quantity and calorie per 100 g if kcal > 100000
foodkcal %>%
  filter(kcal > 100000) %>%
  group_by(description) %>%
  summarise(
    code = mean(code, na.rm = TRUE),
    mean_cons_q = mean(cons_q, na.rm = TRUE),
    mean_kcal100g = mean(kcal100g, na.rm = TRUE),
    .groups = "drop"
  )
##Suggesting that more careful data cleaning might be useful. Leave this for now. 

# Tabulate items where kcal == 0 and cons_q != 0
foodkcal %>%
  filter(kcal == 0, cons_q != 0) %>%
  group_by(description) %>%
  summarise(
    code = mean(code, na.rm = TRUE),
    n = n(),
  ) %>%
  arrange(desc(n))

# Drop cigarettes and cigars
foodkcal <- foodkcal %>%
  filter(!(code %in% c(22011, 22021)))

# Sum the kcal at household level and per capita
kcal<- foodkcal %>%
  group_by(hhid) %>%
  mutate(kcal_total = sum(kcal, na.rm = TRUE)) %>%
  mutate(kcalpc = (kcal_total / (365 / 12)) / hhsize)

# Round kcalpc to integer. 
kcal$kcalpc <- round(kcal$kcalpc, 0)

# Create hhtag which =1 only one time per household. 
kcal <- kcal %>%
  group_by(hhid) %>%
  mutate(hhtag = row_number() == 1) %>%
  ungroup()

# Calculate cutoffs for trimming
cutoffs <- kcal %>%
  filter(hhtag) %>%
  summarise(
    p1 = quantile(kcalpc, 0.01, na.rm = TRUE),
    p99 = quantile(kcalpc, 0.99, na.rm = TRUE)
  )

# Trim top/bottom 1% of kcalpc.
kcal <- kcal %>%
  filter(kcalpc >= cutoffs$p1, kcalpc <= cutoffs$p99)

# Descriptive statistics of kcalpc
skim(kcal$kcalpc)%>%
  yank("numeric")

# Sort by household id and kcal
kcal <-  kcal %>%
  arrange(hhid, desc(kcal))

# Collapse data and getting total kcalpc (mean) and total cons and cons_q by HHs.  
## Note kcalpc is already total as calculated above.
kcal <- kcal %>%
  group_by(hhid) %>%
  summarise(
    kcalpc = mean(kcalpc, na.rm = TRUE),
  )

#Checking final per capita calorie (kcalpc).
skim(kcal$kcalpc)%>%
  yank("numeric")

# labelling 
label(kcal$kcalpc) <- "Daily kilocalories consumed (per capita)"

write_rds(kcal, file.path(data_temp, "kcal.rds"), compress = "gz")
```


###	Regional (NUTS 2) Poverty Lines Estimation
This section estimates regional poverty lines using the Cost of Basic Needs (CBN) method to derive the implicit spatial price index: $\widetilde{SPI_g} = \frac{z_g}{E[z_g]}$ in {@eq-ph-nfood}.

The CBN poverty line is calculated separately for each NUTS 2 region, following the steps below, and the results are saved together in a combined dataset.

- **Step 1**: Calculation of the food poverty line  
- **Step 2**: Calculation of non-food allowances


```{r}
#Load the consumption aggregate data.
ca_nominal_t2 <- readRDS(file.path(data_in, "ca_nominal_t2.rds"))

ca_nominal_t2 <- ca_nominal_t2 %>%
  select(
    hhid,
    starts_with("mpce"),
    hhsize,
    urban,
    location,
    region,
    nuts2,
    mregion,
    starts_with("w")
  )    

ca_kcal <- mg(ca_nominal_t2, kcal, key = c("hhid"))
#Drop trimmed observations (merge=1) for calorie outliers.
ca_kcal <- ca_kcal %>% filter(merge != 1)
ca_kcal <- ca_kcal %>%
  select(-merge)
```

This is the list of the NUTS 2 locations: 

```{r}
# Loop over nuts2
nuts2_list <- unique(ca_nominal_t2$nuts2)
print(nuts2_list)
```

The loop for calculating poverty lines across NUTS 2 regions begins below:

```{r}
#Note run the following loop altogether. 

for (m in nuts2_list) {
  nuts2_data <- ca_kcal %>% filter(nuts2 == m)
  
  nuts2_data <- nuts2_data %>%
    mutate(ukcal = (mpcet_food / (365 / 12)) / kcalpc)
  
  vars_to_drop <- grep("_dec$", names( nuts2_data ), value = TRUE)
  nuts2_data  <-  nuts2_data [ , !(names(nuts2_data) %in% vars_to_drop)]
  
  nuts2_data <- nuts2_data%>%
    arrange(hhid) %>%
    mutate(mpcet_dec = ntile(mpcet, 10))

  #Step 1: Getting Food Poverty lines

    # Food poverty line 
  er <- 2200
  
  # Reference group for mean ukcal 
  
  mukcal <- nuts2_data %>%
    filter(mpcet_dec %in% 1:3) %>%
    summarise(mean_ukcal = mean(ukcal, na.rm = TRUE)) %>%
    pull(mean_ukcal)
  
  nuts2_data <- nuts2_data %>%
    mutate(fpl = round(er * mukcal * (365 / 12)),0)  %>%
    mutate(n_fpl = sum(!is.na(ukcal[mpcet_dec %in% 1:3])))  %>%
    mutate(n_smp = sum(!is.na(ukcal)))
  
  label(nuts2_data$n_fpl) <- "N obs used for food pline"
  label(nuts2_data$n_smp) <- "N obs"          
  label(nuts2_data$fpl) <- "Food PL (PLN/person/month)"
  
  skim(nuts2_data$fpl)%>%
  yank("numeric")

  #Step 2: Getting non-food allowances
  width <- 0.10
  nuts2_data <- nuts2_data %>%
    mutate(
      bot = fpl * (1 - width),
      top = fpl * (1 + width),
      botk = er * (1 - width),
      topk = er * (1 + width)
    )
  
  nuts2_data <- nuts2_data %>%
    mutate(mpcet_nf = mpcet - mpcet_food)
  
  # NFA lower bound / upper 
  
  lbnfa <- nuts2_data %>%
    filter(mpcet >= bot & mpcet <= top) %>%
    summarise(median_mpcet_nf = median(mpcet_nf, na.rm = TRUE)) %>%
    pull(median_mpcet_nf)
  
  ubnfa <- nuts2_data %>%
    filter(mpcet_food >= bot & mpcet_food <= top & kcalpc >= botk & kcalpc <= topk) %>%
    summarise(median_mpcet_nf = median(mpcet_nf, na.rm = TRUE)) %>%
    pull(median_mpcet_nf)
  
  n_ubpl <- nuts2_data %>%
    filter(mpcet_food >= bot & mpcet_food <= top & kcalpc >= botk & kcalpc <= topk) %>%
    summarise(count_ubpl = sum(!is.na(mpcet_nf))) %>%
    pull(count_ubpl)
  
  n_lbpl <- nuts2_data %>%
    filter(mpcet >= bot & mpcet <= top) %>%
    summarise(count_lbpl = sum(!is.na(mpcet_nf))) %>%
    pull(count_lbpl)
  
  nuts2_data <- nuts2_data %>%
    mutate(
      lbpl = fpl + lbnfa,
      ubpl = fpl + ubnfa,
      n_lbpl = n_lbpl,
      n_ubpl = n_ubpl
    )
  
  nuts2_data <- nuts2_data %>%
    mutate(w_z = ifelse(mpcet_food >= bot & mpcet_food <= top & kcalpc >= botk & kcalpc <= topk, mpcet_food / mpcet, NA))
  
  nuts2_summary <- nuts2_data %>%
    group_by(nuts2) %>%
    summarise(
      fpl = mean(fpl, na.rm = TRUE),
      lbnfa = mean(lbnfa, na.rm = TRUE),
      ubnfa = mean(ubnfa, na.rm = TRUE),
      lbpl = mean(lbpl, na.rm = TRUE),
      ubpl = mean(ubpl, na.rm = TRUE),
      n_fpl = mean(n_fpl, na.rm = TRUE),
      n_lbpl = mean(n_lbpl, na.rm = TRUE),
      n_ubpl = mean(n_ubpl, na.rm = TRUE),              
      w_z = mean(w_z, na.rm = TRUE)
    )
  
write_rds(nuts2_summary, file.path(data_temp, paste0("poverty_line_nominal_nuts2_", m, ".rds")), compress = "gz")
}


# Append all nuts2 files
overall_poverty <- bind_rows(lapply(nuts2_list, function(m) {
  readRDS(file.path(data_temp, paste0("poverty_line_nominal_nuts2_", m, ".rds")))
}))

#loop end here

write_rds(overall_poverty, file.path(data_temp,"poverty_line_nominal_nuts2.rds"), compress = "gz")
```

###	Six Steps to Calculate the Full-Coverage Deflator
Follow these six steps to estimate the full-coverage deflator.

#### Step 1
Calculate the **household-level** food Paasche index $P_h^{food}$ in @eq-full-coverage. The formula is already defined in @eq-fpaasche. 

```{r, results='hide'}
### Food component of full-coverage index (hh level)

#Load food item data 
food_expenditure_item <- readRDS(file.path(data_temp, "food_expenditure_item.rds")) %>%
  rename(foodq = cons_q) %>%
  mutate(up = foodexp / foodq)  # Compute mean price at household level

# Compute mean price at NATIONAL level
up0 <- food_expenditure_item %>%
  group_by(code) %>%
  summarise(up0 = mean(up, na.rm = TRUE), .groups = "drop")

# Load household expenditure data
ca_nominal_t2 <- readRDS(file.path(data_in, "ca_nominal_t2.rds"))

# Merge expenditure data (`m:1 hhid`)
food_expenditure_item <- mg(food_expenditure_item , ca_nominal_t2, key=c("hhid"), drop_merge = TRUE) 

# Compute food budget share
food_expenditure_item <- food_expenditure_item %>%
  mutate(w_f = foodexp / mpcet)

var_label(food_expenditure_item$w_f) <- "Food item budget share of total expenditure"

# Merge price data (up0)
food_expenditure_item <- mg(food_expenditure_item, up0, key=c("code"), drop_merge = TRUE)

food_expenditure_item <- food_expenditure_item %>%
  mutate(inv_spi_food_full = w_f * (up0 / up))

# Collapse food components to household level
spi_food_full <- food_expenditure_item %>%
  group_by(hhid) %>%
  summarise(inv_spi_food_full = sum(inv_spi_food_full, na.rm = TRUE),
            region = first(region),
            nuts2 = first(nuts2), .groups = "drop") %>%
  mutate(spi_food_full = 1 / inv_spi_food_full)  # Compute SPI food full

var_label(spi_food_full$spi_food_full) <- "SPI food at household level"

write_rds(spi_food_full, file.path(data_temp, "spi_food_full.rds"), compress = "gz")
```


#### Step 2
Compute the Paasche spatial food price index at NUTS 2 (group**$g$** of ${SPI}_g^{food}$ in @eq-ph-nfood). Note that this is not the same as the household-level food Paasche in @eq-fpaasche.

```{r, results='hide'}
#Compute SPI food (nuts2 level)
food_expenditure_item <- readRDS(file.path(data_temp, "food_expenditure_item.rds")) %>%
  rename(foodq = cons_q) %>%
  mutate(up = foodexp / foodq)   # Mean price at household level

up0 <- food_expenditure_item %>%
  group_by(code) %>%
  summarise(up0 = mean(up, na.rm = TRUE), .groups = "drop")

up_r <- food_expenditure_item %>%
  group_by(code, nuts2) %>%
  summarise(up_r = mean(up, na.rm = TRUE), .groups = "drop")

food_expenditure_item <- mg(food_expenditure_item, ca_nominal_t2, key=c("hhid"))
food_expenditure_item <- food_expenditure_item %>%
  filter(merge==3) %>%
  select(-merge) %>%
  mutate(w_f = foodexp / mpcet)

var_label(food_expenditure_item$w_f) <- "Food item budget share of total expenditure"

# Collapse budget share by nuts2 and item
nuts2_budget_shares <- food_expenditure_item %>%
  group_by(nuts2, code) %>%
  summarise(w_f = mean(w_f, na.rm = TRUE), region = first(region), .groups = "drop")

nuts2_budget_shares <- mg(nuts2_budget_shares,up0, key=c("code"), drop_merge = TRUE)
nuts2_budget_shares <- mg(nuts2_budget_shares,up_r, key=c("code", "nuts2"), drop_merge = TRUE)


# Compute SPI food at regional level
nuts2_budget_shares <- nuts2_budget_shares %>%
  mutate(inv_spi_food = w_f * (up0 / up_r))

spi_food_nuts2 <-  nuts2_budget_shares %>%
  group_by(nuts2) %>%
  summarise(
    inv_spi_food = sum(inv_spi_food, na.rm = TRUE),
    region = first(region),
    .groups = "drop"
  ) %>%
  mutate(
    spi_food = 1 / inv_spi_food,  # Compute SPI food
  )
```
```{r}
skim(spi_food_nuts2$spi_food)%>%
  yank("numeric")

# Generate summary table for counties
nuts2_summary <- spi_food_nuts2 %>%
  group_by(nuts2) %>%
  summarise(
    mean_spi_food = mean(spi_food, na.rm = TRUE),
  )
print(nuts2_summary)

saveRDS(spi_food_nuts2, file.path(data_temp, "spi_food_nuts2.rds"))
```


#### Step 3
Compute $w_z$, the average food budget share of the bottom deciles of the distribution of per capita expenditure.

```{r}
# Non Food Component of full-coverage index
# INDIRECT METHOD 

#Load data
ca_nominal_t2 <- readRDS(file.path(data_in, "ca_nominal_t2.rds")) 
kcal <- readRDS(file.path(data_temp, "kcal.rds"))

ca_nominal_t2 <- ca_nominal_t2%>%
  select(hhid, quarter, starts_with("mpce"), hhsize, region, nuts2)

w_z_data <- mg(ca_nominal_t2, kcal, key=c("hhid")) %>%
  filter(merge==3) %>%
  select(-merge) 
##Drop unmatched cases (merge=1) for households with too low or high calorie.

w_z_data <- w_z_data %>% 
  arrange(mpcet)  %>%
  mutate(mpcet_dec = ntile(mpcet, 10)) 

w_z_data <- w_z_data %>% 
  filter(mpcet_dec %in% c(1, 2, 3)) %>%
  summarise(
    mpcet_food = sum(mpcet_food, na.rm = TRUE),
    mpcet = sum(mpcet, na.rm = TRUE)
  ) %>%
  mutate(w_z = mpcet_food / mpcet)

w_z_value <- mean(w_z_data$w_z, na.rm = TRUE)
cat("w_z is:", w_z_value, "\n")

w_z_data <- w_z_data %>%
  mutate(id_w = 1)  

write_rds(w_z_data, file.path(data_temp, "w_z.rds"), compress = "gz")
```

#### Step 4
Compute the group level poverty lines using the CBN method in order to derive the implicit deflator $\widetilde{SPI_g} =  \frac{z_g}{E[z_g]}$ in @eq-ph-nfood.

```{r}
#  Compute spi_implicit - ratio of poverty lines and avg PL (UB)

nuts2_deflators <- readRDS(file.path(data_temp, "poverty_line_nominal_nuts2.rds"))
spi_implicit <- nuts2_deflators %>%
  mutate(spi_implicit = ubpl / mean(ubpl, na.rm = TRUE))

write_rds(spi_implicit, file.path(data_temp, "spi_implicit.rds"), compress = "gz")
```


#### Step 5
Compute the $P_h(g)^{nfood}$ in @eq-full-coverage.

```{r, results='hide'}
# Compute non food component for indirect method (spi_nfood)
spi_food_nuts2 <- readRDS(file.path(data_temp, "spi_food_nuts2.rds")) %>%
  mutate(id_w = 1)  
w_z <- readRDS(file.path(data_temp, "w_z.rds"))
spi_implicit <- readRDS(file.path(data_temp, "spi_implicit.rds"))

spi_food_nuts2 <- mg(spi_food_nuts2, w_z, key = c("id_w"), drop_merge = TRUE) 

spi_food_nuts2 <- mg(spi_food_nuts2, spi_implicit, key = c("nuts2"), drop_merge = TRUE)

spi_nfood_i <- spi_food_nuts2 %>%
  mutate(inv_spi_nfood = (1 / (1 - w_z)) * spi_implicit - (w_z / (1 - w_z)) * spi_food)

write_rds(spi_nfood_i, file.path(data_temp, "spi_nfood_i.rds"), compress = "gz")
```


#### Step 6
Compute the full-coverage index in @eq-full-coverage.

```{r, results='hide'}
# Load the elements of SPI full-coverage index
spi_full <- readRDS(file.path(data_temp, "spi_food_full.rds"))
spi_nfood_i <- readRDS(file.path(data_temp, "spi_nfood_i.rds"))
spi_full <- mg(spi_full, spi_nfood_i, key = "nuts2",drop_merge = TRUE)


# Compute full-coverage indices
spi_full <- spi_full %>%
  mutate(
    spi_full_indirect = 1 / (inv_spi_food_full + (1 - w_z) * (1 / inv_spi_nfood)),
  )
```
```{r}
# Verify computed variables
summary(spi_full$spi_full_indirect)

# Keep only relevant columns
spi_full <- spi_full %>%
  select(hhid, nuts2, w_z, starts_with("inv_"), starts_with("spi_full"))

# Reorder columns
spi_full <- spi_full %>%
  relocate(hhid, nuts2, w_z, starts_with("inv_"), starts_with("spi_full"))

# Save the cleaned dataset
write_rds(spi_full, file.path(data_temp, "spi_full.rds"), compress = "gz")
```

## Results

Finally, let's bring together all of the estimated price deflators.

### Consolidation of estimated spatial price deflators

```{r,results='hide'}
# Bring all indices together 
# Load the Paasche extended dataset
paasche_extended <- readRDS(file.path(data_temp, "paasche_extended.rds"))

# Merge with spi_food_full
spi_food_full <- readRDS(file.path(data_temp, "spi_food_full.rds"))
paasche_extended <- mg(paasche_extended, spi_food_full, key = "hhid", drop_merge = TRUE)

# Merge with spi_full
spi_full <- readRDS(file.path(data_temp, "spi_full.rds"))
paasche_extended <- mg(paasche_extended, spi_full, key = "hhid", drop_merge = TRUE) 

# Merge with spi_full
spi_full <- readRDS(file.path(data_temp, "spi_full.rds"))
paasche_extended <- mg(paasche_extended, spi_full, key = "hhid", drop_merge = TRUE) 

# Merge with hh_info
weight <- readRDS(file.path(data_in, "hh_info.rds"))
weight <- weight %>% select(hhid,popw,hhw)
paasche_extended <- mg(paasche_extended, weight, key = "hhid", drop_merge = TRUE) 


# Ensure numeric conversion
paasche_extended <- paasche_extended %>%
  mutate(
    spi_food = as.numeric(unlist(spi_food)),
    spi_food_rent = as.numeric(unlist(spi_food_rent))
  )

# Rename labels (for clarity in output)
label_descriptions <- c(
  "Food Paasche - hh level",
  "Extended Paasche (food+ rent) - hh level",
  "Full-coverage SPI - Food component",
  "Full-coverage SPI - Indirect method"
)

names(label_descriptions) <- c(
  "spi_food", "spi_food_rent", 
  "spi_food_full", "spi_full_indirect"
)

# Save the household level spatial price indexes
write_rds(paasche_extended, file.path(data_out, "spi_all_hh.rds"), compress = "gz")
```


```{r}
# Final quick look
fin_check<- paasche_extended %>%
  select(spi_food,spi_food_rent,spi_full_indirect) 
skim(fin_check)%>%
  yank("numeric")
```





The list of the variable names and price indexes estimated at the NUTS 2 levels are shown in the table below.

| Variable        | Price index                    |
|-----------------|--------------------------------|
| spi_food        | Food Paasche                   |
| spi_food_rent   | Extended Paasche (food+ rent)  |
| spi_full        | Full-coverage Price Index      |
: Price Index  {.striped .hover}




The list of variables re-scaled is shown in the next table: 

| Variable        | Price index                            |
|-----------------|----------------------------------------|
| spi_food_r      | Rescaled Food Paasche                  |
| spi_food_rent_r | Rescaled Extended Paasche (food+ rent) |
| spi_full_r      | Rescaled Full-coverage Price Index     |
: Rescaled Price Index  {.striped .hover}



```{r}
# Generate tables to be presented. 
# Collapse (mean) by NUTS2 and rescale
table_nuts2 <- paasche_extended %>%
  group_by(nuts2) %>%
  summarise(
    spi_food = weighted.mean(spi_food, popw, na.rm = TRUE),
    spi_food_rent = weighted.mean(spi_food_rent, popw, na.rm = TRUE),
    spi_full = weighted.mean(spi_full_indirect, popw, na.rm = TRUE)
  ) %>%
  mutate(
    spi_food_r = spi_food / mean(spi_food, na.rm = TRUE),
    spi_food_rent_r = spi_food_rent / mean(spi_food_rent, na.rm = TRUE),
    spi_full_r = spi_full / mean(spi_full, na.rm = TRUE)
  )

# Get value labels for NUTS2
nuts2_labels <- val_labels(table_nuts2$nuts2)
nuts2_label_map <- setNames(names(nuts2_labels), nuts2_labels)

# Add label columns
table_nuts2 <- table_nuts2 %>%
  mutate(
    nuts2_label = nuts2_label_map[as.character(nuts2)],
    nuts2_code_label = paste0(nuts2, ": ", nuts2_label)
  ) %>%
  relocate(nuts2, nuts2_label, nuts2_code_label) 


# Collapse (mean) by macro region and rescale
table_mregion <- paasche_extended %>%
  group_by(mregion) %>%
  summarise(
    spi_food = weighted.mean(spi_food, popw, na.rm = TRUE),
    spi_food_rent = weighted.mean(spi_food_rent, popw, na.rm = TRUE),
    spi_full = weighted.mean(spi_full_indirect, popw, na.rm = TRUE)
  ) %>%
  mutate(
    spi_food_r = spi_food / mean(spi_food, na.rm = TRUE),
    spi_food_rent_r = spi_food_rent / mean(spi_food_rent, na.rm = TRUE),
    spi_full_r = spi_full / mean(spi_full, na.rm = TRUE)
  )

# Extract label vector
mregion_labels <- val_labels(table_mregion$mregion)
# Create named vector: name = label, value = number
mregion_label_map <- setNames(names(mregion_labels), mregion_labels)
# Add two new columns: code + label and label only
table_mregion <- table_mregion %>%
  mutate(
    mregion_code_label = paste0(mregion, ": ", mregion_label_map[as.character(mregion)]),
    mregion_label = mregion_label_map[as.character(mregion)]
  ) %>%
relocate(mregion, mregion_label, mregion_code_label) 
```


### Saving in Excel and displaying table results 
```{r}
excel_path <- file.path(tables, "spi_R.xlsx")
wb <- createWorkbook()

# Write to Excel
addWorksheet(wb, "NUTS2")
spi_nuts2 <- labelled::remove_labels(table_nuts2)
writeData(wb, sheet = "NUTS2", table_nuts2)
addWorksheet(wb, "Macro region")
writeData(wb, sheet = "Macro region", table_mregion)

# Save the Excel file
saveWorkbook(wb, file = excel_path, overwrite = TRUE)
cat("An Excel file is saved at:", excel_path, "\n")

## Preparing concise table output:
# Round numeric columns to 3 digits
spi_nuts2_rounded <- table_nuts2 %>%
  select(nuts2_code_label, spi_food_r, spi_food_rent_r, spi_full_r) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3)))
```
Table: NUTS 2 Spatial Price Indexes {#tbl-spi-nuts2}
```{r , echo=FALSE, results='asis'}
# Create the table
knitr::kable(spi_nuts2_rounded)
```
```{r}
# Round numeric columns to 3 digits
mregion_labels <- val_labels(table_mregion$mregion) #save vlabel
spi_mregion_rounded <- table_mregion %>%
  select(mregion_code_label,spi_food_r,	spi_food_rent_r,spi_full_r) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3)))
```
Table: Macro Region Spatial Price Indexes {#tbl-spi-macro}

```{r , echo=FALSE, results='asis'}
# Create the table
knitr::kable(spi_mregion_rounded)
```

### Appendix Tables: Full Result Tables
Table: NUTS 2 Full Spatial Price Indexes {#tbl-spi-nuts2-full}
```{r , echo=FALSE, results='asis'}
## Preparing concise table output:
# Round numeric columns to 3 digits
table_nuts2_rounded <- table_nuts2 %>%
  select(nuts2_code_label, spi_food, spi_food_rent, spi_full, spi_food_r, spi_food_rent_r, spi_full_r) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3)))
# Create the table
knitr::kable(table_nuts2_rounded)
```
Table: Macro Region Full Spatial Price Indexes {#tbl-spi-macro-full}
```{r , echo=FALSE, results='asis'}
## Preparing concise table output:
# Round numeric columns to 3 digits
table_mregion_rounded <- table_mregion %>%
  select(mregion_code_label, spi_food, spi_food_rent, spi_full, spi_food_r, spi_food_rent_r, spi_full_r) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3)))
# Create the table
knitr::kable(table_mregion_rounded)
```

